---
title: "ex34_code"
author: "James Grecian"
date: "18/02/2022"
output: html_document
---

```{r knitr_setup, include = FALSE}
knitr::opts_chunk$set(echo = T, message = FALSE, warning = FALSE)
```

```{r load libraries, message = FALSE, warning = FALSE}
require(tidyverse, quietly = TRUE)
require(foieGras, quietly = TRUE)
require(rnaturalearth)
require(sf)
require(patchwork)
```

### Load harp seal SRDL data, fit SSM to four animals with a regular 12 hour interval and extract predicted locations.

```{r fit ssm, message = FALSE, results = 'hide'}

locs <- readRDS("../data/ex33.rds")
fit <- fit_ssm(locs, model = "crw", time.step = 12, vmax = 4)

# Extract fitted values from model
plocs <- grab(fit, "predicted", as_sf = F)

```

### Visualise predicted harp seal paths on a map

```{r map the tracks, fig.height = 6}
# Generate a global shapefile and a simple plot
world <- ne_countries(scale = "medium", returnclass = "sf")

# Create an sf version of the locs data with a WGS84 projection and add to the plot
plocs_sf <- plocs %>% st_as_sf(coords = c('lon', 'lat')) %>% st_set_crs(4326)

# To generate a plot with less distortion first define a projection i.e. Lambert Azimuthal Equal Area
prj = "+proj=laea +lat_0=60 +lon_0=-50 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

p1 <- ggplot() +
  theme_bw() +
  geom_sf(aes(), data = world) +
  geom_sf(aes(colour = id), data = plocs_sf, size = 0.75, show.legend = "point") +
  scale_color_viridis_d(option = "cividis") +
  coord_sf(xlim = c(-2500000, 2500000), ylim = c(-2500000, 2500000), crs = prj, expand = T) +
  scale_x_continuous(breaks = seq(from = -130, to = 30, by = 10)) +
  facet_wrap(~id)
print(p1)

```

### Simulate 50 animal tracks from the fitted crw process model

```{r simulate}
trs <- simfit(fit,
              what = "predicted",
              reps = 50,
              grad = readRDS(system.file("extdata/grad.RDS", package = "foieGras")),
              beta = c(-300, 300))
```

### Filter the simulations based on similarity to the original path and retain top 10%

```{r filter}
trs_flt <- sim_filter(trs,
                      keep = .1,
                      flag = 2)
```

### Plot comparison of filtered and unfiltered paths

```{r comparison plot, fig.width = 9, fig.height = 6, fig.align = 'center'}

trs <- trs %>% unnest(cols = c(sims)) %>% st_as_sf(coords = c('lon', 'lat')) %>% st_set_crs(4326)
trs_flt <- trs_flt %>% unnest(cols = c(sims)) %>% st_as_sf(coords = c('lon', 'lat')) %>% st_set_crs(4326)

p2 <- ggplot() +
  theme_bw() +
  geom_sf(aes(), data = world) +
  geom_sf(aes(colour = factor(rep)), data = trs, size = 0.5) +
  scale_color_viridis_d(option = "cividis") +
  coord_sf(xlim = c(-4000000, 5500000), ylim = c(-4500000, 7500000), crs = prj, expand = T) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 20)) +
  facet_wrap(~id, ncol = 2) +
  theme(legend.position="none") +
  ggtitle("All simulations")

    
p3 <- ggplot() +
  theme_bw() +
  geom_sf(aes(), data = world) +
  geom_sf(aes(colour = factor(rep)), size = 0.5, data = trs_flt) +
  scale_color_viridis_d(option = "cividis") +
  coord_sf(xlim = c(-4000000, 5500000), ylim = c(-4500000, 7500000), crs = prj, expand = T) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 20)) +
  facet_wrap(~id, ncol = 2) +
  theme(legend.position="none") +
  ggtitle("Top 10%")

print(p2 + p3)

```

is it going to be good enough? hp6-756-19 isn't a great fit...

could do the move to land stuff from Josh London...
