---
title: "Appendix S1: R code for Application 3.1"
author: "Jonsen et al"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load required R packages
```{r load pkgs, eval=FALSE, message=FALSE}
require(tidyverse)
require(patchwork)
require(sf)
require(foieGras)
```

### Fit `rw` and `crw` SSM's with 12-h `time.step` to southern elephant seal example data
```{r fit SSMs, eval=FALSE, message=FALSE}
## access data from foieGras
data(sese2, package = "foieGras")

## use only 1 seal track
sese2 <- subset(sese2, id == "ct36-E-09")

fit.rw <- fit_ssm(sese2, model = "rw", time.step = 12)
fit.crw <- fit_ssm(sese2, model = "crw", time.step = 12)
```

### Calculate One-Step-Ahead (Prediction) residuals for both model fits
```{r calc osars, eval=FALSE, message=FALSE}

res.rw <- osar(fit.rw)
res.crw <- osar(fit.crw)
```

## Create Figure 1
### Panel A - model fits to observed locations
```{r panel a, eval=FALSE, message=FALSE}

## Use sf package to project locations & convert SSM-estimated locations to lines
dat <- grab(fit.rw, "d", as_sf = TRUE) %>% 
  sf::st_transform("+proj=stere +lon_0=90 +units=km")

loc.rw <- grab(fit.rw, "f", as_sf = TRUE) %>%
  sf::st_transform("+proj=stere +lon_0=90 +units=km") 

line.rw <- loc.rw %>%
  dplyr::group_by(id) %>%
  dplyr::summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTILINESTRING")

loc.crw <- grab(fit.crw, "f", as_sf = TRUE) %>%
  sf::st_transform("+proj=stere +lon_0=90 +units=km") 

line.crw <- loc.crw %>%
  dplyr::group_by(id) %>%
  dplyr::summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTILINESTRING")

line.ssm <- dplyr::bind_rows(line.rw, line.crw) %>%
  dplyr::mutate(SSM = c("rw","crw")) %>%
  dplyr::select(id, SSM, geometry)

## contruct inset zoom box
zbox <- structure(list(id = c(1,1,1,1,1), 
                       lon = c(107, 110.25, 110.25, 107, 107), 
                       lat = c(-66.5, -66.5, -64.9, -64.9, -66.5)), 
                  class = "data.frame", row.names = c(NA, -5L))

## project zoom box
prj <- sf::st_crs(line.ssm)

zbox_sf <- zbox %>%
  sf::st_as_sf(coords = c("lon","lat"), crs = 4326) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(geometry = sf::st_combine(geometry)) %>%
  sf::st_cast("POLYGON")

## get zoom box bounds
bb_zbox <- sf::st_bbox(zbox_sf)

## Plot SSM fits to observed locations
## main map
pA <- ggplot() +
  geom_sf(data = dat, 
             col = "orange", 
             size = 1.25) +
  geom_sf(data = line.ssm,
            aes(colour = SSM)) +
  geom_sf(data = zbox_sf, size = 0.5, colour = "black", fill = NA) +
  scale_colour_manual(values = c("firebrick", "dodgerblue")) +
  labs(x = element_blank(), y = element_blank()) +
  theme_minimal() +
  theme(axis.text = element_text(size = 8),
        legend.key.size = unit(c(6, 3), "mm"),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 8),
        legend.position = c(0.2,0.2))

## inset map
pAsub1 <- ggplot() +
  geom_sf(data = dat, 
             col = "orange", 
             size = 1) +
  geom_sf(data = line.ssm,
            aes(colour = SSM)) +
  coord_sf(xlim = c(bb_zbox[1], bb_zbox[3]), 
           ylim = c(bb_zbox[2], bb_zbox[4]), 
           crs = prj, 
           expand = FALSE) +
  scale_colour_manual(values = c("firebrick", "dodgerblue"), guide = "none") +
  labs(x = element_blank(), y = element_blank()) +
  theme_minimal() +
  theme(axis.text = element_blank(),
        panel.background = element_rect(colour = "white"),
        panel.border = element_rect(fill=NA, size=0.5))

## position inset on main map
pAsub1 <- ggplotGrob(pAsub1)

foo <- sf::st_bbox(dat)

pAA <- pA + 
  annotation_custom(grob = pAsub1, 
                             xmin = foo$xmin + (foo$xmax - foo$xmin) * 0.35, 
                             xmax = foo$xmin + (foo$xmax - foo$xmin) * 1.30, 
                             ymin = foo$ymin + (foo$ymax - foo$ymin) * 0.35, 
                             ymax = foo$ymin + (foo$ymax - foo$ymin) * 1.05)
```


### Panel B - plot time-series of prediction residuals for `rw` model
```{r panel b, eval=FALSE, message=FALSE}

## use `foieGras::plot.osar()`; modify with additional ggplot2 functions
pB <- plot(res.rw, "ts") + 
  xlab(element_blank()) + 
  ylab("Residuals") +
  theme(axis.text.x = element_text(size=7, angle=45), 
        axis.title.y = element_text(size = 8),
        strip.text.y = element_blank(),
        axis.title.x = element_blank())
```


### Panel C - plot autocorrelation functions of prediction residuals for `rw` model
```{r panel c, eval=FALSE, message=FALSE}

## use `foieGras::plot.osar()`; modify with additional ggplot2 functions
pC <- plot(res.rw, "acf") + 
  theme(strip.text.y = element_blank(),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 8))
```


### Panel D - plot time-series of prediction residuals for `crw` model
```{r panel d, eval=FALSE, message=FALSE}

## use `foieGras::plot.osar()`; modify with additional ggplot2 functions
pD <- plot(res.crw, "ts") + 
  xlab(element_blank()) + 
  ylab(element_blank()) +
  theme(axis.text.x = element_text(size=7, angle=45), 
        strip.text.y = element_blank(),
        axis.title.x = element_blank())
```


### Panel E - plot autocorrelation functions of prediction residuals for `crw` model
```{r panel e, eval=FALSE, message=FALSE}

## use `foieGras::plot.osar()`; modify with additional ggplot2 functions
pE <- plot(res.crw, "acf") + 
  ylab(element_blank()) +
  theme(strip.text.y = element_blank(),
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 8))
```


### Arrange panels & render figure 1 using the `patchwork` package
```{r figure 1, eval=FALSE, fig.width=7, fig.height=6.25}

pAA / (pB | pD) / (pC | pE) +
  patchwork::plot_layout(widths = c(3,3,3),
              heights = c(3,1,1)) +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 9))
```


