# 2 | `foieGras` Overview

```{r part_2_premble, include=FALSE}
#### Code for part 2: foieGras Overview ####
```
The workflow for `foieGras` is deliberately simple, with much of the usual track data processing checks and formatting handled automatically. The main functions are listed in Table \ref{tab:func_table_kable}. When fitting a model, `foieGras` automatically detects the type of tracking data location quality classes designations that are typical of Argos data and that can be added to the data by the researcher for other types of track data. Based on the location quality classes and other, optional information on observation errors contained in the data, `foieGras` chooses an appropriate measurement error model for each observation. This capability allows for combinations of different tracking data types, e.g., Argos and GPS, in a single input data frame and to be fit in a single state-space model. 

```{r func_table, echo=FALSE,  fig.width=5, fig.height=6, message=FALSE, warning=FALSE, cache=TRUE}
func_table <- data.frame(
  Function = c("fit_mpm", "fit_ssm", "fmap", "grab", "osar", "plot.fG_mpm", "plot.fG_osar", "plot.fG_ssm"),
  Description = c("Fit a Move Persistence Model to location data",
                  "Fit a State-Space Model to location data",
                  "Plot fitted/predicted locations on a map with or without a defined projection",
                  "Extract fitted/predicted/observed locations from a foieGras model, with or without projection information",
                  "Estimate One-Step-Ahead Residuals from a foieGras SSM",
                  "Plot move persistence estimates as 1-D or 2-D (along track) time-series",
                  "Plot One-Step-Ahead Residuals from a foieGras SSM",
                  "Visualise the fit of a foieGras SSM to data"
  )
)

```

```{r func_table_kable, echo=FALSE, fig.width=5, fig.height=6, message=FALSE, warning=FALSE, cache=TRUE, purl=FALSE}
kable(func_table, 
      format = table_out_format, 
      caption="Main functions for the R package \\texttt{foieGras}", 
      booktabs = TRUE) %>%
  column_spec(1, monospace = TRUE) %>%
  column_spec(2, latex_column_spec = ">{\\\\hangindent=2em}p{25em}" ) %>%
  
kable_styling(full_width = FALSE)
```

## 2.1 | Data preparation
Animal tracking data, consisting of a time-series of location coordinates, can be read into R as a data frame using standard functions such as `read.csv`. The canonical data format for Argos tracks consists of a data frame with 5 columns corresponding to the following named variables: `id` (individual id), `date` (date and time), `lc` (location class), `lon` (longitude), `lat` (latitude). Optionally, an additional 3 columns, `smaj` (semi-major axis), `smin` (semi-minor axis), `eor` (ellipse orientation), providing Argos error ellipse information may be included.

Other types of track data can be accommodated, for example, by including the `lc` column where all `lc = "G"` for GPS data. In this case, measurement error in the GPS locations is assumed to have a standard deviation of 0.1 x Argos class 3 locations (approximately 30 m). Other types of track data can be considered in a similar manner (see the package vignette for further details).

## 2.2 | State-space model fitting - `fit_ssm`
State-space models are fit using `fit_ssm`. There are a large number of options that can be set in `fit_ssm` (see Suppl for details). We focus only the essential options here:  

- `data` the input data structured as described in **2.1** 
- `vmax` a maximum threshold speed (ms^-1^) to help identify potential outlier locations 
- `model` the process model to be used 
- `time.step` the prediction time interval (h)

The function first invokes an automated data processing stage where the following occurs: 1) data type (Argos Least-Squares, Argos Kalman Filter/Smoother, GPS, or General (e.g., processed light-level geolocations, acoustic telemetry, coded VHF telemetry) is determined; 2) datetimes are converted to POSIXt format, chronological order is ensured, and duplicate datetime records are removed; 3) observations occurring less than `min.dt` seconds after a prior observation are removed; 4) a speed filter [@Freitas:2008] is used to identify potential outlier locations; 5) locations are projected from spherical lon-lat coordinates to planar x,y coordinates in km. 

The function then fits a state-space model to the processed data, where the process model (currently, either a continuous-time `rw` or a continuous-time `crw`) is specified by the user and the measurement model(s) are selected automatically [see @Jonsen:2020 for model details]. The model is fit by numerical optimization of the likelihood using either the `optim` or `nlminb` R function. The R package `TMB`, Template Model Builder [@Kristensen:2016], is used to compute the gradient function in C++ via reverse-mode auto-differentiation and the Laplace Approximation is used to integrate out the latent states (random effects). Fits to a single versus multiple individuals are handled automatically, with sequential SSM fits occurring in the latter case. No hierarchical or pooled estimation among individuals is currently available. 

`fit_ssm` returns a `foieGras` fit object (a nested data frame with class `fG_ssm`). The outer data frame lists the individual id(s), basic convergence information and a list with class `ssm`. This list contains dense information on the model parameter and state estimates, predictions, processed data, optimizer results, and other diagnostic and contextual information. Users can extract a simple data frame of SSM fitted (location estimates corresponding to the, typically irregular, observation times) or predicted values (locations predicted at regular `time.step` intervals) using the `grab` function. 

## 2.3 | Model checking and visualisation
Before using fitted or predicted locations, a model fit should be checked and visualised to confirm that underlying assumptions have been met. Goodness-of-fit of an SSM can be assessed using One-Step-Ahead residuals [@Thygesen:2017], which are estimated using the `osar` function. The `osar` function uses the `oneStepPredict` function from the `TMB` R package.


## 2.4 | Behavioural estimation


## 2.5 | Simulation
