# 2 | `foieGras` overview

```{r part_2_premble, include=FALSE}
#### Code for part 2: foieGras Overview ####
```
The workflow for `foieGras` is deliberately simple, with many of the usual track data processing checks and formatting handled automatically. Here we outline the main aspects of the `foieGras` package. The packages's main functions are listed in Table \ref{tab:func_table_kable}. 

```{r func_table, echo=FALSE,  fig.width=5, fig.height=6, message=FALSE, warning=FALSE, cache=TRUE}
func_table <- data.frame(
  Function = c("fit_ssm", "fit_mpm", "grab", "osar", "map", "sim", "simfit", "sim_filter", "route_path", "plot.ssm_df", "plot.osar"),
  Description = c("Fit a State-Space Model to location data",
                  "Fit a Move Persistence Model to location data",
                  "Extract fitted/predicted/observed locations from a foieGras model, with or without projection information",
                  "Estimate One-Step-Ahead Residuals from a foieGras SSM",
                  "Map fitted/predicted locations with or without a defined projection",
                  "Simulate individual animal tracks with Argos LS or KF errors",
                  "Simulate animal tracks from \`ssm_df\` fit objects",
                  "Filter tracks simulated with `simfit` according to similarity criteria",
                  "Reroute path so estimated locations are off land",
                  "Plot the fit of a foieGras SSM to data",
                  "Plot One-Step-Ahead Residuals from a foieGras SSM"
  )
)

```

```{r func_table_kable, echo=FALSE, fig.width=5, fig.height=6, message=FALSE, warning=FALSE, cache=TRUE, purl=FALSE}
kable(func_table, 
      format = table_out_format, 
      caption="Main functions for the R package \\texttt{foieGras}", 
      booktabs = TRUE) %>%
  column_spec(1, monospace = TRUE) %>%
  column_spec(2, latex_column_spec = ">{\\\\hangindent=2em}p{25em}" ) %>%
  
kable_styling(full_width = FALSE)
```

## 2.1 | Data preparation
Animal tracking data, consisting of a time-series of location coordinates, can be read into R as a data frame using standard functions such as `read.csv`. The canonical data format for Argos tracks consists of a data frame with 5 columns corresponding to the following named variables: `id` (individual id), `date` (date and time), `lc` (location class), `lon` (longitude), `lat` (latitude). Optionally, an additional 3 columns, `smaj` (semi-major axis), `smin` (semi-minor axis), `eor` (ellipse orientation), providing Argos error ellipse information may be included.

Other types of location data can be accommodated, for example, by including the `lc` column where all `lc = "G"` for GPS data. In this case, measurement error in the GPS locations is assumed to have a standard deviation of 0.1 x Argos class 3 locations (approximately 30 m). If location standard errors exist, these can be added by appending the columns `lonerr` and `laterr` to the data (see the Overview vignette for further details).

## 2.2 | State-space model fitting - `fit_ssm`
State-space models are fit using `fit_ssm`. When fitting a SSM to location data, the type of data is automatically detected from the location quality class designations that are typical of Argos data and that can be added to the data by the researcher for other types of location data. Based on the location quality classes and optional information on measurement errors contained in the data, an appropriate measurement error model is selected for each observation [@Jonsen:2020]. This capability can allow different tracking data types, such as Argos and GPS, to be combined in a single input data frame and to be fit in a single state-space model. 

There are a large number of arguments that can be set in `fit_ssm`, and these are explained in the documentation. We focus only the essential arguments here:  

- `data` the input data structured as described in **2.1** 
- `model` the process model to be used 
- `time.step` the prediction time interval (h)

The function first invokes an automated data processing stage where the following occurs: 1) data type (Argos Least-Squares, Argos Kalman Filter/Smoother, GPS, or General (e.g., processed light-level geolocations, acoustic telemetry, coded VHF telemetry) is determined; 2) date-times are converted to POSIXt format, chronological order is ensured, and duplicate date-time records are removed; 3) observations occurring less than `min.dt` seconds after a prior observation are removed (default: 60 s); 4) a speed filter [`sda` from the `trip` R package; @Sumner:2009] is used to identify potential extreme locations to be ignored by the SSM; 5) locations are projected from spherical lon,lat coordinates to global Mercator x,y coordinates in km. 

The function then fits a state-space model to the processed data, where the process model (one of `rw`, `crw`, or `mp`) is specified by the user via the `model` argument, and the measurement model(s) are selected automatically. The model is fit by numerical optimization of the likelihood using either of the standard R optimizers, `optim` or `nlminb`. The R package `TMB`, Template Model Builder [@Kristensen:2016], is used to compute the gradient function in C++ via reverse-mode auto-differentiation and the Laplace Approximation is used to integrate out the latent states (random effects). Fits to a single versus multiple individuals are handled automatically, with sequential SSM fits occurring in the latter case. No hierarchical or pooled estimation among individuals is currently available. 

`fit_ssm` returns a `ssm_df` fit object (a nested data frame with class `ssm_df`). The outer data frame lists the individual id(s), basic convergence information and a list with class `ssm`. This list contains dense information on the estimated parameters and states, predictions, processed data, optimizer results, and other diagnostic and contextual information. Users can extract a simple data frame of SSM fitted (location estimates corresponding to the observation times) or predicted values (locations predicted at regular `time.step` intervals) using the `grab` function. Parameter estimates, AIC and other model fit information can be viewed in tabular form using the `summary` function.

## 2.3 | Behavioural estimation - `fit_ssm`, `fit_mpm`
Move persistence, an index of along-track movement behaviour, can be estimated as a continuous-valued (0 - 1), time-varying latent variable that represents changes in movement pattern based on autocorrelation in speed and direction [@Auger-Methe:2017,@Jonsen:2019]. There are two approaches in `foieGras` for estimating move persistence. The first is to use `fit_ssm` with `model = 'mp'`, which fits a continuous-time move persistence model in state-space form and thereby simultaneously estimates true locations and move persistence from the error-prone telemetry data. This approach is most appropriate for fitting to irregularly-timed and error-prone Argos data as both aspects are taken into account explicitly. The second is to use `fit_mpm`, which can take as input either location data or SSM-estimated locations from an `ssm_df` fit object. This approach is generally more appropriate when the data have minimal measurement error (e.g., GPS locations), or when  time-regularization is desired. We illustrate both approaches in Application 3.2. 

## 2.4 | Model checking and visualization - `osar`, `plot`, `map`
Before using fitted or predicted locations, a `fit_ssm` model fit should be checked and visualized to confirm that the model adequately describes the data. There is no simple way to calculate residuals for latent variable models that have non-finite state-spaces and that may be nonlinear, but they can be computed based on iterative forecasts of the model [@Thygesen:2017]. The `osar` function computes one-step-ahead (prediction) residuals via the  `oneStepPredict` function from the `TMB` R package to make this as efficient as possible. A set of residuals are calculated for the `x` and `y` values corresponding to the fitted values from the SSM and returned as an `osar` object. A generic `plot` (`plot.osar`) method provides an easy way to visualize the `osar` residuals as time-series plots, quantile-quantile plots, or autocorrelation functions.

State-space model fits to data can also be visualised by using the generic `plot` (`plot.ssm_df`) function on an `ssm_df` fit object. Options exist to plot fitted or predicted values along with observations as either paired, 1-D time-series, as 2-D tracks with 95% confidence intervals or ellipses, by using the argument `type = 1` and `type = 2`, respectively. These plots provide a rapid check on SSM fits to data. Additionally, when the fitted SSM is the move persistence model (i.e., `model = 'mp'`), 1-D time-series (`type = 3`) or 2-D track plots (`type = 4`) of move persistence can be viewed.

Additionally, a `fit_ssm` model fit can be mapped using the `map` function for single or multiple individuals. By default, `map` uses the coastline data from the `rnaturalearth` R package [@rnaturalearth] at medium or high resolution, if the `rnaturalearthhires` [@rnaturalearthhires] R package is installed, but can also use tiled maps for finer-scale detail, if the `rosm` [@rosm] and `ggspatial` [@ggspatial] R packages are installed. Mapping aesthetics (e.g., plot symbols, sizes, colours, fills) can be customized via the `aes` argument and use of the `aes_lst` function. See code in SI for examples.

All `foieGras` visualizations draw on the `ggplot2` R package [@ggplot2], with multi-panel plots also using the `patchwork` R package [@patchwork], and generally can be modified through additive calls in the usual `ggplot2` manner. See code in SI for examples.

## 2.5 | Simulation - `sim`, `simfit`, `sim_filter`
Track simulation can be a helpful, yet informal, way of evaluating the degree to which statistical movement models capture essential features of animal movement data [@Michelot:2017]. The `sim` function can simulate a variety of movement process, including the `rw`, `crw`, and `mp` process models, as well as simple multiple movement state switching processes. Simulations from different process models can be used to evaluate the robustness of SSM location and move persistence estimates to model misspecification. An example is provided in the SI.

Simulation is also used frequently in habitat usage modelling to provide a measure of habitat availability [@Aarts:2012] by generating a source of 'background' points representing a null model of the distribution of foraging animals in the absence of external drivers [@Phillips:2009;@Raymond:2015;@Hindell:2020]. The `simfit` function extracts movement parameters from a `ssm_df` fit object and simulates a user defined number of random tracks of the same duration from these parameters. The argument `cpf = TRUE` allows the user to simulate central place foragers by ensuring that the simulated tracks start and end at approximately the same location. It is also possible to constrain movements to remain mostly in water via a potential function [@Preisler:2013], using included gradient rasters and the `grad` and `beta` arguments. These are illustrated in the code for Application 3.3.

The choice of null points can have a large impact on the performance of habitat suitability models [@Phillips:2009; @Lobo:2010], and so the `sim_filter` function provides a tool to filter the simulated tracks based on their similarity to the original path. The filtering is based on one of two metrics that capture the difference in the net displacement and bearing between the two paths (see `similarity_flag` for more detail). This metric is motivated by the 'flag value' described in @Hazen:2017. The user can also specify the quantile of flag values to retain; i.e. `keep = 0.25` (the default) will return a `simfit` object containing those simulated tracks with flag values in the top 25% of values calculated for the input `simfit` object.

## 2.6 | Path rerouting - `route_path`
As the SSMs implemented in `foieGras` have no information about potential barriers to animal movement it is possible for locations to be estimated in implausible locations, such as on land for marine species. To overcome this, `foieGras` makes use of the `pathroutr` R package [@pathroutr] to efficiently re-route locations from land back to water by using visibility graphs [@Jan:2014]. The `route_path` function can be applied to either a `fit_ssm` model fit (`ssm_df` object) or the simulations generated by `simfit`. When the input is an `ssm_df` object the re-routed path can be appended to the object for visualisation and use in subsequent analyses. When the input is a `simfit` object the locations within the simulation are replaced with the re-routed paths. We illustrate how the latter can be achieved in Application 3.3. 
