# 3 | Applications

```{r part_3_premble, include=FALSE}
#### Code for part 3: Applications ####
```
We illustrate the main capabilities of `foieGras` through a set of applications that are for demonstration purposes and not intended as a comprehensive guide for conducting analyses with `foieGras`. Complete code and data for reproducing the applications and for gaining a deeper understanding of `foieGras` functions are provided in the Supplementary Information.

## 3.1 | SSM validation with prediction residuals
We use a sub-adult male southern elephant seal (*Mirounga leonina*) track included as example data in `foieGras` (`sese2`, id: ct36-E-09), sourced from from the Australian Integrated Marine Observing System (IMOS; data publicly available via [imos.aodn.org.au](http://imos.aodn.org.au)) deployments at Iles Kerguelen in collaboration with the French IPEV and SNO-MEMO programmes. The data are temporally irregular Argos Least-Squares based locations, 74 \% of which are in the poorest location quality classes: A and B. We fit both the `rw` and `crw` models using `fit_ssm` with a speed filter threshold (`vmax`) of 4 ms^-1^ and a 12-h time step. We calculate prediction residuals using `osar`, and then use the generic `plot` method for `osar` residuals to assess and compare the model fits (Fig. \ref{fig:ex1}).

```{r ex31, echo=FALSE, message=FALSE, results='markup', fig.width=7, fig.height=6.25, cache=TRUE, fig.cap="\\label{fig:ex1} State-space model fits to a southern elephant seal track (a), and diagnostic plots for assessing goodness-of-fit of the `rw` (b - prediction residual time-series; d - prediction residual autocorrelation) and `crw` (c,e) state-space models. All residual plots generated using the `plot.osar` function."}

load("../data/ex31.RData")

dat <- grab(fit.rw[1,], "d", as_sf = TRUE) |> 
  sf::st_transform("+proj=stere +lon_0=90 +units=km")
loc.rw <- grab(fit.rw[1,], "f", as_sf = TRUE) |>
  sf::st_transform("+proj=stere +lon_0=90 +units=km") 
line.rw <- loc.rw |>
  dplyr::group_by(id) |>
  dplyr::summarise(do_union = FALSE, .groups = "drop") |>
  sf::st_cast("MULTILINESTRING")
loc.crw <- grab(fit.crw[1,], "f", as_sf = TRUE) |>
  sf::st_transform("+proj=stere +lon_0=90 +units=km") 
line.crw <- loc.crw |>
  dplyr::group_by(id) |>
  dplyr::summarise(do_union = FALSE, .groups = "drop") |>
  sf::st_cast("MULTILINESTRING")

line.ssm <- bind_rows(line.rw, line.crw) |>
  mutate(SSM = c("rw","crw")) |>
  select(id, SSM, geometry)

pA <- ggplot() +
  geom_sf(data = dat, 
             col = "dodgerblue", 
             size = 1.5) +
  geom_sf(data = line.ssm,
            aes(colour = SSM)) +
  scale_colour_manual(values = c("orange", "firebrick")) +
  labs(x = element_blank(), y = element_blank()) +
  theme_minimal() +
  theme(axis.text = element_text(size = 6),
        legend.key.size = unit(c(5, 4), "mm"),
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6),
        legend.position = c(0.25,0.2))

dist <- data.frame(dist = 
                     sf::st_distance(loc.rw, loc.crw, by_element = TRUE)
                   ) |>
  mutate(dist = as.numeric(dist))

pB <- ggplot(dist) +
  geom_histogram(aes(dist),
                 fill = grey(0.6),
                 col = "white",
                 bins = 20) +
  labs(x = "Distance (km)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 8),
        axis.text = element_text(size = 6),
        plot.background = element_rect(fill = "white", colour = NA))

pAB <- pA + 
  patchwork::inset_element(pB, 0.5, 0.55, 1, 1, ignore_tag = TRUE) +
  patchwork::plot_annotation(tag_levels = "a")

p1 <- plot(res.rw, "ts") + 
  xlab(element_blank()) + 
  ylab("Residuals") +
  theme(axis.text.x = element_text(size=6, angle=45), 
        axis.title.y = element_text(size = 8),
        strip.text.y = element_blank(),
        axis.title.x = element_blank())

p2 <- plot(res.rw, "acf") + 
  theme(strip.text.y = element_blank(),
        axis.text = element_text(size = 6),
        axis.title = element_text(size = 8))

p3 <- plot(res.crw, "ts") + 
  ylab(element_blank()) + 
  xlab(element_blank()) + 
  theme(axis.text.x = element_text(size=6, angle=45), 
        strip.text.y = element_blank(),
        axis.title.x = element_blank())

p4 <- plot(res.crw, "acf") + 
  ylab(element_blank()) + 
  theme(strip.text.y = element_blank(),
        axis.text = element_text(size = 6),
        axis.title = element_text(size = 8))

pA / (p1 | p3) / (p2 | p4) +
  plot_layout(widths = c(2,3,3),
              heights = c(2,1,1)) +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 9))
```

The plots of predicted states on top of the observations suggest both models yield similar fits (Fig. \ref{fig:ex1}a; orange vs red lines). However, corresponding predicted locations from the two models differ by a median 6.62 km (range: 0.02, 53.02 km), and there are marked trends in the time-series of residuals for the `rw` model fit (Fig. \ref{fig:ex1}b) with significantly positive autocorrelation in both the x and y directions (Fig. \ref{fig:ex1}d). The `crw` prediction residuals show little trend through time and have relatively little autocorrelation (Fig. \ref{fig:ex1}c,e), implying that the `crw` process model provides a better fit to the data.

## 3.2 | Inferring move persistence from Argos and GPS data
### ***Argos data - southern elephant seals***
Drawing on additional IMOS tracking data from Iles Kerguelen, we infer move persistence, $\gamma_t$, along four southern elephant seal tracks. We fitted the `mp` SSM with a 12-h prediction interval with `fit_ssm`, using the Argos Kalman filter-derived error ellipse information provided with each observation [@Jonsen:2020]. The `mp` SSM simultaneously estimates locations and $\gamma_t$, and their uncertainties. We then assess how $\gamma_t$ changes along the seals' tracks to infer regions where the seals spend disproportionately more or less time during their foraging trips.  


```{r ex32.1, echo=FALSE, message=FALSE, results='markup', fig.width=6, fig.height=7, cache=TRUE, fig.cap="\\label{fig:ex32.1} Inferred move persistence, $\\gamma_t$, time-series for four southern elephant seals (a-d; grey envelopes are 95 % CI's, note differing date ranges on the x axes), and along their 2-D tracks (e; track labels, a-d, correspond to the time-series plots). Locations associated with low $\\gamma_t$ (purple) are indicative of slow, undirected movements, whereas high $\\gamma_t$ (yellow) is indicative of faster, directed movements. Blue circles highlight bouts of spatially constrained low $\\gamma_t$."}
load("../data/ex32_fit.RData")

p1 <- plot(fit, 
           type = 3, 
           pages = 1, 
           ncol=2, 
           ask=0,
           normalise = TRUE,
           group = TRUE,
           pal = "Plasma") & 
  ylim(0, 1) &
#  xlim(ISOdate(2021,01,01,tz="UTC"), ISOdate(2021,10,05,tz="UTC")) &
  theme(legend.position = "none",
        axis.text = element_text(size = 7))

my.aes <- aes_lst(conf = FALSE, 
                  line = TRUE)
my.aes$df$size[1] <- 1.5
my.aes$df$size[3] <- 0.1
my.aes$df$col[3] <- grey(0.3)
my.aes$df$fill[5] <- grey(0.4)
my.aes$df$fill[6] <- grey(0.6)

m1 <- map(fit, 
          what = "p", 
          aes = my.aes,
          crs = "+proj=stere +lon_0=95 +units=km +datum=WGS84 +no_defs",
          normalise = TRUE, 
          group = TRUE,
          silent = TRUE) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  ggspatial::annotation_scale(height = unit(1.25, "mm"), 
                              aes(width_hint = 0.2,
                                  location = "br",
                                  text_col = "white")) + 
  theme(legend.position = c(0.95,0.5), 
        legend.direction = "vertical",
        legend.key.width = unit(4, "mm"),
        legend.key.height = unit(7, "mm"),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 7),
        axis.text = element_text(size = 7),
        panel.grid = element_line(colour = "white"))
  

sf_locs <- grab(fit, what = "p", as_sf = TRUE) %>%
  sf::st_transform(., crs = m1$coordinates$crs)
bb <- sf::st_bbox(sf_locs)

label.df <- data.frame(tag = c("a","b","c","d"), 
                        x = c(0.75, 0.9, 0.19, 0.53) * 
                         (bb["xmax"] - bb["xmin"]) + bb["xmin"],
                        y = c(0.32, 0.75, 0.2, 0.25) * 
                         (bb["ymax"] - bb["ymin"]) + bb["ymin"])
circle.df <- data.frame(x = c(0.25, 0.365, 0.485, 0.71) *
                          (bb["xmax"] - bb["xmin"]) + bb["xmin"],
                        y = c(0.11, 0.19, 0.24, 0.26) * 
                         (bb["ymax"] - bb["ymin"]) + bb["ymin"],
                        r = c(125,125,125,145)
                       )
  
m1 <- m1 + 
  geom_text(data = label.df, 
            aes(x, y, label=tag),
            size = 3) +
  ggforce::geom_circle(
             aes(x0 = circle.df$x, y0 = circle.df$y, r = circle.df$r),
             fill = NA,
             size = 0.8,
             colour = "dodgerblue")

(p1 | m1) +
  plot_layout(nrow = 2, 
              heights = c(2, 5), 
              guides = "keep") +
  plot_annotation(tag_levels = "a") &
  theme(plot.title = element_blank(),
        plot.tag = element_text(size = 9, colour = "black"))
```


The three southern elephant seals on foraging trips to the Antarctic shelf region all engaged in spatially constrained bouts of low move persistence while in the shelf region (Fig. \ref{fig:ex32.1} a,c,d; blue circles in e). Without additional data it is unclear exactly what these bouts of low horizontal move persistence represent. They could result from area-restricted search and foraging within dense prey aggregations, physical constraints of dense ice on horizontal movements, haulout (resting) on sea-ice, or some combination of these. Conversely, the seal on a pelagic foraging trip engaged in slower, more meandering movements with less spatially constrained bouts of lower move persistence (Fig. \ref{fig:ex32.1} b, e). This general movement pattern may be consistent with searching for suitable foraging resources within the highly variable eddy fields between the Subantarctic and Polar Fronts [@Jonsen:2019].  

### ***GPS data - little penguins***
To illustrate how move persistence can be estimated from other types of animal tracking data, we use four little penguin (*Eudyptula minor*) GPS tracks from daily foraging trips during the chick-rearing period from Montague Island, NSW, Australia, and described in @Phillips:2021. The data are temporally irregular GPS locations, with high frequency sampling (15 s on average) intermittently disrupted by the birds' diving behaviour, and are assumed to have minimal measurement error. We fitted the `crw` SSM to the GPS data to predict temporally regular locations at 5-min intervals, and assumed consistently small bivariate normal location measurement errors (ie. $\pm$ 10 m sd). We then used `fit_mpm` to estimate $\gamma_t$ from these regularised locations.  


```{r ex32.2, echo=FALSE, message=FALSE, results='markup', fig.width=7, fig.height=4, cache=TRUE, fig.cap="\\label{fig:ex32.2} Inferred move persistence, $\\gamma_t$, 1-D time-series (a-d; grey envelopes are 95 % CI's) and along little penguin GPS tracks (e)."}
load("../data/ex32_lipe.RData")

p <- plot(fmp, ncol = 2, pages = 1, ask = FALSE, pal = "Plasma") &
  theme(legend.position = "none",
        plot.title = element_blank(),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank())

peng.ssm_sf <- foieGras::grab(fit, "p", as_sf = TRUE)
bb <- sf::st_bbox(peng.ssm_sf)
my.aes <- aes_lst(conf = F, line = T, mp_pal = hcl.colors(n=100, "Plasma"))
my.aes$df$size[1] <- 1

m <- map(
  fit,
  fmp,
  what = "p",
  aes = my.aes,
  map_type = "cartolight",
  zoomin = 1,
  ext.rng = c(0.3, 0.1),
  normalise = FALSE,
  silent = TRUE,
  alpha = 0.75
) +
  ggspatial::annotation_scale(height = unit(0.15, "cm"), 
                              aes(location = "br")) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  scale_x_continuous(breaks = pretty(seq(150.13, 150.26, l = 4), n = 3)) +
  scale_y_continuous(breaks = pretty(seq(-36.5, -36.24, l = 5), n = 4)) +
  theme(legend.position = c(0.88,0.84),
        legend.key.width = unit(4, "mm"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        axis.text = element_text(size = 6),
        panel.grid = element_line(colour = "black"))

label.df <- data.frame(tag = c("a","b","c","d"), 
                        x = c(0.1, 0.84, 0, 0.86) * 
                         (bb["xmax"] - bb["xmin"]) + bb["xmin"],
                        y = c(0.9, 0.225, 0.45, 0.55) * 
                         (bb["ymax"] - bb["ymin"]) + bb["ymin"])
m <- m + 
  geom_text(data = label.df, aes(x,y,label=tag), size = 3)

## plot movement persistence 1-D time-series and map along 2-D tracks
(p | m) +
  plot_layout(widths = c(3, 2), guides = "keep") +
  plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 9))
```

The little penguin GPS tracks did not exhibit strong contrast in move persistence, with $\gamma_t$ declining below 0.5 for only two of the birds (Fig. \ref{fig:ex32.2}a-d). Nonetheless, the move persistence estimates highlight change in movement pattern over the course of the penguins' daily foraging trips (Fig. \ref{fig:ex32.2}e). The penguins departed Montague Is. with relatively fast movements, three directed southward (Fig. \ref{fig:ex32.2} tracks b-d in e) and one less directed and remaining close to the island (track a), before slowing down and engaging in meandering movements (orange - red in Fig. \ref{fig:ex32.2}e). The spatially diffuse bouts of low move persistence within the penguin tracks may reflect the fine-scale patchiness of their forage-fish prey with search and prey-capture occurring both within and among discrete neighbouring prey aggregations [@Carroll:2017]. **Lach/Gemma could we overlay SVM-inferred prey captures for these birds? (G123f02, G124m10, G126m05, L013m01)**  


## 3.3 | Simulating tracks from `foieGras` model fits

To illustrate how to simulate tracks from `foieGras` model fits we use a juvenile harp seal (*Pagophilus groenlandicus*) tracked from the Gulf of St Lawrence, Canada, and described in @Grecian:2022. The data are temporally irregular Argos locations including error ellipse information. We fit the `crw` model using `fit_ssm` with a 4 ms^-1^ speed filter threshold (`vmax`) and a 12-h prediction interval (`time.step`).

We simulate 50 animal movement paths from the `crw` process model using `simfit`, and apply a potential function using the `grad` and `beta` arguments to constrain the simulated paths to largely remain in water. These tracks are then filtered based on their similarity to the original path using `sim_filter` and the top 10% retained (`keep = 0.1`)(Fig. \ref{fig:ex33}a,b). As the potential function does not guarantee all locations remain off land, we re-route any remaining simulated locations from land back to water using `route_path` (Fig. \ref{fig:ex33}c). In combination, these functions provide a pragmatic, non-statistical method to generate and objectively filter pseudo-tracks for use in movement or habitat modelling applications.

```{r ex33, echo = FALSE, message = FALSE, results = 'markup', fig.width = 9, fig.height = 6, cache = TRUE, fig.cap = "\\label{fig:ex33} Simulating (a) 100 movement paths from a correlated random walk process model; (b) filtering those tracks to select the top 10% based on their similarity to the original SSM-predicted track (red); and (c) re-routing simulated locations on land (orange) back to ocean (blue). The orange box in (b) indicates region magnified in (c). SSM-predicted track (red) overlaid in all panels for context."}
# load track data
locs <- readRDS("../data/ex33.rds")
# load gradient rasters for simfit
load(system.file("extdata/grad.rda", package = "foieGras"))

fit <- fit_ssm(locs %>% filter(id == "hp6-749-19"), 
               model = "crw", 
               time.step = 12, 
               vmax = 4, 
               control = ssm_control(verbose = 0))

set.seed(10000)
trs <- foieGras::simfit(fit,
              what = "predicted",
              reps = 100,
              grad = grad,
              beta = c(-350, -350))

trs_flt <- foieGras::sim_filter(trs,
                      keep = 0.1,
                      flag = 2)

trs_flt_rr <- foieGras::route_path(trs_flt,
                                map_scale = 50,
                                buffer = 20000)

world_shp <- rnaturalearth::ne_countries(scale = 50, returnclass = "sf")

# To generate a plot with less distortion first define a projection i.e. Lambert Azimuthal Equal Area
prj <- "+proj=laea +lat_0=60 +lon_0=-50 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

trs_sf <- trs %>% 
  tidyr::unnest(cols = c(sims)) %>% 
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  sf::st_transform(crs = prj) %>%
  group_by(id, rep) %>%
  summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTIPOINT")

trs_flt_sf <- trs_flt %>% 
  tidyr::unnest(cols = c(sims)) %>% 
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  sf::st_transform(crs = prj) %>%
  group_by(id, rep) %>%
  summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTIPOINT")

trs_flt_rr_sf <- trs_flt_rr %>% 
  tidyr::unnest(cols = c(sims)) %>% 
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  sf::st_transform(crs = prj) %>%
  group_by(id, rep) %>%
  summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTIPOINT")

plocs_sf <- foieGras::grab(fit, "predicted", as_sf = TRUE) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTIPOINT")

pline_sf <- foieGras::grab(fit, "predicted", as_sf = TRUE) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(do_union = FALSE, .groups = "drop") %>%
  sf::st_cast("MULTILINESTRING")

bb_trs <- sf::st_bbox(trs_sf)  

p1 <- ggplot() +
  theme_minimal() +
  geom_sf(data = world_shp, colour = NA, fill = grey(0.6)) +
  geom_sf(data = trs_sf, size = 0.2, colour = "dodgerblue") +
  geom_sf(data = plocs_sf, size = 0.5, colour = "firebrick") +
  coord_sf(xlim = c(bb_trs[1], bb_trs[3]), 
           ylim = c(bb_trs[2], bb_trs[4]), 
           crs = prj, 
           expand = TRUE) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 20)) +
  scale_y_continuous(breaks = seq(from = 0, to = 90, by = 15)) +
  ggspatial::annotation_scale(height = unit(1.5, "mm"), aes(location = "br"))

## create box to magnify in subsequent maps
zbox <- structure(list(id = c(1,1,1,1,1), 
                          lon = c(-53, -68, -68, -53, -53), 
                          lat = c(45, 45, 51.25, 51.25, 45)), 
                  class = "data.frame", row.names = c(NA, -5L))
zbox_sf <- zbox %>%
  sf::st_as_sf(coords = c("lon","lat"), crs = 4326) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(geometry = sf::st_combine(geometry)) %>%
  sf::st_cast("POLYGON")

bb_zbox <- sf::st_bbox(zbox_sf)  

p2 <- ggplot() +
  theme_minimal() +
  geom_sf(data = world_shp, colour = NA, fill = grey(0.6)) +
  geom_sf(size = 0.2, data = trs_flt_sf, colour = "dodgerblue") +
  geom_sf(data = plocs_sf, size = 0.4, colour = "firebrick") +
  geom_sf(data = zbox_sf, size = 0.75, colour = "orange", fill = NA) +
  coord_sf(xlim = c(bb_trs[1], bb_trs[3]), 
           ylim = c(bb_trs[2], bb_trs[4]), 
           crs = prj, 
           expand = TRUE) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 20)) +
  scale_y_continuous(breaks = seq(from = 0, to = 90, by = 15)) +
  ggspatial::annotation_scale(height = unit(1.5, "mm"), aes(location = "br"))

world_shp <- rnaturalearth::ne_countries(scale = 10, returnclass = "sf")

p3 <- ggplot() +
  theme_minimal() +
  geom_sf(data = world_shp, colour = NA, fill = grey(0.6)) +
  geom_sf(data = trs_flt_sf, colour = "orange", size = 0.8) +
  geom_sf(data = trs_flt_rr_sf, colour = "dodgerblue", size = 0.8) +
  geom_sf(data = pline_sf, size = 0.2, colour = "firebrick") +
  geom_sf(data = plocs_sf, size = 0.4, colour = "firebrick") +
  coord_sf(xlim = c(bb_zbox[1], bb_zbox[3]), 
           ylim = c(bb_zbox[2], bb_zbox[4]), 
           crs = prj, 
           expand = FALSE) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 5)) +
  scale_y_continuous(breaks = seq(from = 0, to = 90, by = 5)) +
  ggspatial::annotation_scale(height = unit(1.5, "mm"), aes(location = "tl"))

p1 + p2 + p3 + 
  patchwork::plot_layout(ncol = 2) +
  patchwork::plot_annotation(tag_levels = "a") & 
  theme(plot.tag = element_text(size = 10))
```



