# 3 | Examples

```{r part_3_premble, include=FALSE}
#### Code for part 3: Examples ####
```
We illustrate the main capabilities of `foieGras` through a series of examples that are for demonstration purposes and not intended as a comprehensive guide for conducting analyses with `foieGras`. Complete code and data for reproducing the examples and for gaining a deeper understanding of `foieGras` functions are provided as supplements.

## 3.1 | Southern Elephant seal - SSM validation with prediction residuals
We use a sub-adult male southern elephant seal track included in `foieGras` (`sese2`, id: ct36-E-09), sourced from from the Australian Integrated Marine Observing System (IMOS; data publicly available via [imos.aodn.org.au](http://imos.aodn.org.au)) deployments at Iles Kerguelen in collaboration with the French IPEV and SNO-MEMO programmes. The data are temporally irregular Argos Least-Squares based locations, 74 \% of which are in the poorest location quality classes: A and B. We fit both the `rw` and `crw` models using `fit_ssm` with a speed filter threshold (`vmax`) of 4 ms^-1^ and a 12-h time step. We calculate prediction residuals using `osar`, and then use the generic `plot` method for `osar` residuals to assess and compare the model fits (Fig. \ref{fig:ex1}).

```{r ex31, echo=FALSE, message=FALSE, results='markup', fig.width=7, fig.height=6.25, cache=TRUE, fig.cap="\\label{fig:ex1} State-space model fits to a southern elephant seal track (a), and diagnostic plots for assessing `rw` (c - time-series of prediction residuals; e - autocorrelation of prediction residuals) and `crw` (d,f) state-space model goodness-of-fit. All residual plots generated using the `plot.fG_osar` function."}

load("../data/ex31.RData")

dat <- grab(fit.rw[1,], "d", as_sf = TRUE) |> 
  sf::st_transform("+proj=stere +lon_0=90 +units=km")
loc.rw <- grab(fit.rw[1,], "f", as_sf = TRUE) |>
  sf::st_transform("+proj=stere +lon_0=90 +units=km") 
line.rw <- loc.rw |>
  dplyr::group_by(id) |>
  dplyr::summarise(do_union = FALSE) |>
  sf::st_cast("MULTILINESTRING")
loc.crw <- grab(fit.crw[1,], "f", as_sf = TRUE) |>
  sf::st_transform("+proj=stere +lon_0=90 +units=km") 
line.crw <- loc.crw |>
  dplyr::group_by(id) |>
  dplyr::summarise(do_union = FALSE) |>
  sf::st_cast("MULTILINESTRING")

line.ssm <- bind_rows(line.rw, line.crw) |>
  mutate(SSM = c("rw","crw")) |>
  select(id, SSM, geometry)

pA <- ggplot() +
  geom_sf(data = dat, 
             col = "dodgerblue", 
             size = 1.5) +
  geom_sf(data = line.ssm,
            aes(colour = SSM)) +
  scale_colour_manual(values = c("orange", "firebrick")) +
  labs(x = element_blank(), y = element_blank()) +
  theme_minimal() +
  theme(axis.text = element_text(size = 6),
        legend.key.size = unit(c(5, 4), "mm"),
        legend.title = element_text(size = 7),
        legend.text = element_text(size = 6),
        legend.position = c(0.3,0.2))

dist <- data.frame(dist = 
                     sf::st_distance(loc.rw, loc.crw, by_element = TRUE)
                   ) |>
  mutate(dist = as.numeric(dist))

pB <- ggplot(dist) +
  geom_histogram(aes(dist),
                 fill = grey(0.6),
                 col = "white",
                 bins = 20) +
  labs(x = "Distance (km)") +
  theme_minimal() +
  theme(axis.title = element_text(size = 8),
        axis.text = element_text(size = 6),
        plot.background = element_rect(fill = "white", colour = NA))

pAB <- pA + 
  patchwork::inset_element(pB, 0.5, 0.55, 1, 1, ignore_tag = TRUE) +
  patchwork::plot_annotation(tag_levels = "a")

p1 <- plot(res.rw, "ts") + 
  xlab(element_blank()) + 
  ylab("Residuals") +
  theme(axis.text.x = element_text(size=6, angle=45), 
        axis.title.y = element_text(size = 8),
        strip.text.y = element_blank(),
        axis.title.x = element_blank())

p2 <- plot(res.rw, "acf") + 
  theme(strip.text.y = element_blank(),
        axis.text = element_text(size = 6),
        axis.title = element_text(size = 8))

p3 <- plot(res.crw, "ts") + 
  ylab(element_blank()) + 
  xlab(element_blank()) + 
  theme(axis.text.x = element_text(size=6, angle=45), 
        strip.text.y = element_blank(),
        axis.title.x = element_blank())

p4 <- plot(res.crw, "acf") + 
  ylab(element_blank()) + 
  theme(strip.text.y = element_blank(),
        axis.text = element_text(size = 6),
        axis.title = element_text(size = 8))

pA / (p1 | p3) / (p2 | p4) +
  plot_layout(widths = c(2,3,3),
              heights = c(2,1,1)) +
  patchwork::plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 9))
```

The plots of predicted states on top of the observations suggest both models yield similar fits (Fig. \ref{fig:ex1}a; orange vs red lines). However, corresponding predicted locations from the two models differ by a median 6.62 km (range: 0.02, 53.02 km), and there are marked trends in the time-series of residuals for the `rw` model fit (Fig. \ref{fig:ex1}b) with significantly positive autocorrelation in both the x and y directions (Fig. \ref{fig:ex1}d). The `crw` prediction residuals show little trend through time and have relatively little autocorrelation (Fig. \ref{fig:ex1}c,e), implying that the `crw` process model provides a better fit to the data.



## 3.2 | Inferring movement persistence from Argos and GPS data
Drawing on an expanded version of the data used in 3.1, we quality control and infer movement persistence, $\gamma_t$, along four southern elephant seal tracks. We fitted the `mp` SSM with a 24-h prediction interval using `fit_ssm`, assuming bivariate normal measurement errors consistent with Argos Least-Squares-derived locations [@Jonsen:2020]. The `mp` SSM simultaneously estimates locations and $\gamma_t$, and their uncertainties. We then assess how $\gamma_t$ changes along the seals' tracks to infer regions where the seals spend disproportionately more or less time during their foraging trips.

To illustrate how the method can accommodate other types of animal tracking data, we also infer $\gamma_t$ along four little penguin (*Eudyptula minor*) GPS tracks from Montague Island, NSW Australia, described in @Phillips:2021. The data are temporally irregular GPS locations, with high frequency sampling (1 - 2s) periodically disrupted by the birds' diving behaviour, and are assumed to have minimal measurement error. We consequently chose to fit the `crw` SSM to the GPS data and predict temporally regular locations at 5-min intervals, assuming consistently small bivariate normal location measurement errors (ie. $\pm$ 10 m sd). We then used `fit_mpm` to estimate $\gamma_t$ from the SSM-predicted locations.

```{r ex32.1, echo=FALSE, message=FALSE, results='markup', fig.width=8, fig.height=6.5, cache=TRUE, fig.cap="\\label{fig:ex32.1} Inferred move persistence, $\\gamma_t$, 1-D time-series for four southern elephant seals (a-d; grey envelopes are 95 % CI's) and along their 2-D tracks (e; track labels, a-d, correspond to the 1-D time-series plots). Locations associated with low move persistence (purple) are indicative of slow, undirected movements, whereas high move persistence (yellow) is indicative of faster, directed movements."}
load("../data/ex32.RData")

p1 <- plot(fit[1:4,], 
           type = 3, 
           pages = 1, 
           ncol=2, 
           ask=0,
           pal = "Plasma") &
  theme(legend.position = "none",
        axis.text = element_text(size = 7))

my.aes <- aes_lst(conf = FALSE, 
                  line = TRUE,
                  mp_pal = hcl.colors(n=100, "Plasma"))
my.aes$df$fill[5] <- grey(0.8)
my.aes$df$fill[6] <- grey(0.6)

m1 <- map(fit[1:4,], 
          what = "p", 
          aes = my.aes,
          crs = "+proj=stere +lon_0=70 +units=km +datum=WGS84 +no_defs",
          silent = TRUE) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  ggspatial::annotation_scale(height = unit(1.25, "mm"), 
                              aes(width_hint = 0.1,
                                  location = "br")) + 
  theme(legend.position = c(0.85,0.9), 
        legend.direction = "horizontal",
        legend.key.width = unit(7, "mm"),
        legend.key.height = unit(4, "mm"),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 7),
        axis.text = element_text(size = 7),
        panel.grid = element_line(colour = "white"))

sf_locs <- grab(fit, what = "p", as_sf = TRUE) %>%
  sf::st_transform(., crs = m1$coordinates$crs)
bb <- sf::st_bbox(sf_locs)

label.df <- data.frame(tag = c("a","b","c","d"), 
                        x = c(1.03, 0.03, -0.03, 0.51) * 
                         (bb["xmax"] - bb["xmin"]) + bb["xmin"],
                        y = c(0.56, 0.08, 0.68, 0.3) * 
                         (bb["ymax"] - bb["ymin"]) + bb["ymin"])
m1 <- m1 + 
  geom_text(data = label.df, 
            aes(x, y, label=tag),
            size = 4)

(p1 / m1) +
  plot_layout(nrow = 2, 
              heights = c(2, 3), 
              guides = "keep") +
  plot_annotation(tag_levels = "a") &
  theme(plot.title = element_blank(),
        plot.tag = element_text(size = 10, colour = "black"))
```

Movement persistence estimates along the quality-controlled southern elephant seal tracks highlight some fundamental differences in movement pattern among the seals. The two seals engaging in pelagic foraging trips (Fig. \ref{fig:ex32.1}a,c and e) had less contrast in their movements with consistently higher $\gamma_t$ estimates compared to the two seals engaging in trips to the fast-ice on the Antarctic shelf (Fig. \ref{fig:ex32.1}b,d and e). Although $\gamma_t$'s were higher overall for the pelagically foraging seals, they both spent little time making fast, highly directed movements ($\gamma_t \to 1$) relative to the shelf-foraging seals (\ref{fig:ex32.1}a,c vs b,d). This suggests the pelagically-foraging seals may spend considerable time searching for suitable foraging habitat in the highly variable eddy fields between the Subantarctic and Polar Fronts [@Jonsen:2019], whereas foraging habitat may be more predictable for seals travelling rapidly and directly to the Antarctic shelf region. These seals may also haulout periodically on available fast-ice to rest. This behaviour could also contribute to the higher contrast in movement persistence, relative to pelagically-foraging seals who would not have access to fast-ice.

```{r ex32.2, echo=FALSE, message=FALSE, results='markup', fig.width=7, fig.height=4, cache=TRUE, fig.cap="\\label{fig:ex32.2} Inferred move persistence, $\\gamma_t$, 1-D time-series (a-d; grey envelopes are 95 % CI's) and along little penguin GPS tracks (e)."}
load("../data/ex32_lipe.RData")

p <- plot(fmp, ncol = 2, pages = 1, ask = FALSE, pal = "Plasma") &
  theme(legend.position = "none",
        plot.title = element_blank(),
        axis.text = element_text(size = 6),
        panel.grid.minor = element_blank())

peng.ssm_sf <- foieGras::grab(fit, "p", as_sf = TRUE)
bb <- sf::st_bbox(peng.ssm_sf)
my.aes <- aes_lst(conf = F, line = T, mp_pal = hcl.colors(n=100, "Plasma"))
my.aes$df$size[1] <- 1

m <- map(
  fit,
  fmp,
  what = "p",
  aes = my.aes,
  map_type = "cartolight",
  zoomin = 1,
  ext.rng = c(0.3, 0.1),
  normalise = FALSE,
  silent = TRUE,
  alpha = 0.75
) +
  ggspatial::annotation_scale(height = unit(0.15, "cm"), 
                              aes(location = "br")) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  scale_x_continuous(breaks = pretty(seq(150.13, 150.26, l = 4), n = 3)) +
  scale_y_continuous(breaks = pretty(seq(-36.5, -36.24, l = 5), n = 4)) +
  theme(legend.position = c(0.88,0.84),
        legend.key.width = unit(4, "mm"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        axis.text = element_text(size = 6),
        panel.grid = element_line(colour = "black"))

label.df <- data.frame(tag = c("a","b","c","d"), 
                        x = c(0.1, 0.84, 0, 0.86) * 
                         (bb["xmax"] - bb["xmin"]) + bb["xmin"],
                        y = c(0.9, 0.225, 0.45, 0.55) * 
                         (bb["ymax"] - bb["ymin"]) + bb["ymin"])
m <- m + 
  geom_text(data = label.df, aes(x,y,label=tag), size = 3)

## plot movement persistence 1-D time-series and map along 2-D tracks
(p | m) +
  plot_layout(widths = c(3, 2), guides = "keep") +
  plot_annotation(tag_levels = "a") &
  theme(plot.tag = element_text(size = 9))
```

Despite vastly different scales of movement, the time series of little penguin movement persistence estimates were broadly similar to those of the southern elephant seals (Fig. \ref{fig:ex32.2}a-d). The little penguin foraging trips likely reflect the underlying spatial distribution of their forage-fish prey, with spatially diffuse bouts of lower movement persistence potentially indicative of foraging both within and among neighbouring discrete prey patches [@Carroll:2017] (Fig. \ref{fig:ex32.2}e).


## 3.3 | Simulating tracks from `foieGras` model fits

To illustrate how to simulate tracks from `foieGras` model fits we use a juvenile harp seal (*Pagophilus groenlandicus*) tracked from the Gulf of St Lawrence, Canada, and described in @Grecian:2022. The data are temporally irregular Argos locations including error ellipse information. We fit the `crw` model using `fit_ssm` with a 4 ms^-1^ speed filter threshold (`vmax`) and a 12-h prediction interval (`time.step`).

We simulate 50 animal movement paths from the `crw` process model using `simfit`, and apply a potential function using the `grad` and `beta` arguments to constrain the simulated paths to largely remain in water. These tracks are then filtered based on their similarity to the original path using `sim_filter` and the top 10% retained (`keep = 0.1`)(Fig. \ref{fig:ex33}a,b).

As the potential function does not guarantee all locations remain off land, we re-route any remaining simulated locations from land back to water using `route_path` (Fig. \ref{fig:ex33}c).

In combination, these functions provide a simple, pragmatic method to generate and objectively filter pseudo-tracks for use in movement or habitat modelling applications.

```{r ex33, echo = FALSE, message = FALSE, results = 'markup', fig.width = 9, fig.height = 6, cache = TRUE, fig.cap = "\\label{fig:ex33} Simulating (a) 100 movement paths from a correlated random walk process model; (b) filtering those tracks to select the top 10% based on their similarity to the original SSM-predicted track (red); and (c) re-routing simulated locations on land (orange) back to ocean (blue). The orange box in (b) indicates region magnified in (c). SSM-predicted track (red) overlaid in all panels for context."}
# load track data
locs <- readRDS("../data/ex33.rds")
# load gradient rasters for simfit
load(system.file("extdata/grad.rda", package = "foieGras"))

fit <- fit_ssm(locs %>% filter(id == "hp6-749-19"), 
               model = "crw", 
               time.step = 12, 
               vmax = 4, 
               control = ssm_control(verbose = 0))

set.seed(10000)
trs <- foieGras::simfit(fit,
              what = "predicted",
              reps = 100,
              grad = grad,
              beta = c(-350, -350))

trs_flt <- foieGras::sim_filter(trs,
                      keep = 0.1,
                      flag = 2)

trs_flt_rr <- foieGras::route_path(trs_flt,
                                map_scale = 50,
                                buffer = 20000)

world_shp <- rnaturalearth::ne_countries(scale = 50, returnclass = "sf")

# To generate a plot with less distortion first define a projection i.e. Lambert Azimuthal Equal Area
prj <- "+proj=laea +lat_0=60 +lon_0=-50 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"

trs_sf <- trs %>% 
  tidyr::unnest(cols = c(sims)) %>% 
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  sf::st_transform(crs = prj) %>%
  group_by(id, rep) %>%
  summarise(do_union = FALSE) %>%
  sf::st_cast("MULTIPOINT")

trs_flt_sf <- trs_flt %>% 
  tidyr::unnest(cols = c(sims)) %>% 
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  sf::st_transform(crs = prj) %>%
  group_by(id, rep) %>%
  summarise(do_union = FALSE) %>%
  sf::st_cast("MULTIPOINT")

trs_flt_rr_sf <- trs_flt_rr %>% 
  tidyr::unnest(cols = c(sims)) %>% 
  sf::st_as_sf(coords = c('lon', 'lat'), crs = 4326) %>% 
  sf::st_transform(crs = prj) %>%
  group_by(id, rep) %>%
  summarise(do_union = FALSE) %>%
  sf::st_cast("MULTIPOINT")

plocs_sf <- foieGras::grab(fit, "predicted", as_sf = TRUE) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(do_union = FALSE) %>%
  sf::st_cast("MULTIPOINT")

pline_sf <- foieGras::grab(fit, "predicted", as_sf = TRUE) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(do_union = FALSE) %>%
  sf::st_cast("MULTILINESTRING")

bb_trs <- sf::st_bbox(trs_sf)  

p1 <- ggplot() +
  theme_minimal() +
  geom_sf(data = world_shp, colour = NA, fill = grey(0.6)) +
  geom_sf(data = trs_sf, size = 0.2, colour = "dodgerblue") +
  geom_sf(data = plocs_sf, size = 0.5, colour = "firebrick") +
  coord_sf(xlim = c(bb_trs[1], bb_trs[3]), 
           ylim = c(bb_trs[2], bb_trs[4]), 
           crs = prj, 
           expand = TRUE) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 20)) +
  scale_y_continuous(breaks = seq(from = 0, to = 90, by = 15)) +
  ggspatial::annotation_scale(height = unit(1.5, "mm"), aes(location = "br"))

## create box to magnify in subsequent maps
zbox <- structure(list(id = c(1,1,1,1,1), 
                          lon = c(-53, -68, -68, -53, -53), 
                          lat = c(45, 45, 51.25, 51.25, 45)), 
                  class = "data.frame", row.names = c(NA, -5L))
zbox_sf <- zbox %>%
  sf::st_as_sf(coords = c("lon","lat"), crs = 4326) %>%
  sf::st_transform(crs = prj) %>%
  group_by(id) %>%
  summarise(geometry = sf::st_combine(geometry)) %>%
  sf::st_cast("POLYGON")

bb_zbox <- sf::st_bbox(zbox_sf)  

p2 <- ggplot() +
  theme_minimal() +
  geom_sf(data = world_shp, colour = NA, fill = grey(0.6)) +
  geom_sf(size = 0.2, data = trs_flt_sf, colour = "dodgerblue") +
  geom_sf(data = plocs_sf, size = 0.4, colour = "firebrick") +
  geom_sf(data = zbox_sf, size = 0.75, colour = "orange", fill = NA) +
  coord_sf(xlim = c(bb_trs[1], bb_trs[3]), 
           ylim = c(bb_trs[2], bb_trs[4]), 
           crs = prj, 
           expand = TRUE) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 20)) +
  scale_y_continuous(breaks = seq(from = 0, to = 90, by = 15)) +
  ggspatial::annotation_scale(height = unit(1.5, "mm"), aes(location = "br"))

world_shp <- rnaturalearth::ne_countries(scale = 10, returnclass = "sf")

p3 <- ggplot() +
  theme_minimal() +
  geom_sf(data = world_shp, colour = NA, fill = grey(0.6)) +
  geom_sf(data = trs_flt_sf, colour = "orange", size = 0.8) +
  geom_sf(data = trs_flt_rr_sf, colour = "dodgerblue", size = 0.8) +
  geom_sf(data = pline_sf, size = 0.2, colour = "firebrick") +
  geom_sf(data = plocs_sf, size = 0.4, colour = "firebrick") +
  coord_sf(xlim = c(bb_zbox[1], bb_zbox[3]), 
           ylim = c(bb_zbox[2], bb_zbox[4]), 
           crs = prj, 
           expand = FALSE) +
  scale_x_continuous(breaks = seq(from = -180, to = 180, by = 5)) +
  scale_y_continuous(breaks = seq(from = 0, to = 90, by = 5)) +
  ggspatial::annotation_scale(height = unit(1.5, "mm"), aes(location = "tl"))

p1 + p2 + p3 + 
  patchwork::plot_layout(ncol = 2) +
  patchwork::plot_annotation(tag_levels = "a") & 
  theme(plot.tag = element_text(size = 10))
```



