---
title: "`foieGras` an R package for animal movement data: rapid quality control, behavioural estimation and simulation"
header-includes:
  - \usepackage{placeins}
  - \usepackage{fontspec}
  - \setmainfont[Ligatures=TeX, BoldFont={Helvetica Neue Medium}, ItalicFont={Helvetica Neue Light Italic}]{Helvetica Neue Light}
  - \usepackage{sectsty}
  - \sectionfont{\sffamily\large\scshape}
  - \subsectionfont{\sffamily\normalsize\scshape}
compact-title: TRUE
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    includes:
        in_header:
          - preamble.sty
documentclass: "article"
classoption: "11pt"
geometry: left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm,headheight=12pt,a4paper
bibliography: references.bib
csl: methods-in-ecology-and-evolution.csl
spacing: double
---

```{r knitr_setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=TRUE,
                      warning=FALSE,
                      cache=TRUE,
                      autodep=TRUE,
                      fig.path="../figures/",
                      #to handle remaining figures without indices
                      #From:https://github.com/yihui/knitr/issues/505
                      fig.process = function(x) {
                        x2 = sub('-\\d+[.][a-z]+)$', '\\1', x)
                        if (file.rename(x, x2)) x2 else x 
                        }
                      )
```

```{r setup, include=FALSE, cache=FALSE}
#### Setup ####
## all the packages needed for this tutorial are listed here
library(foieGras)
library(mpmm)
library(ggplot2)
library(ggthemes)
library(patchwork)
library(dplyr)
library(lubridate)
library(kableExtra)
library(knitr)
library(docxtools)
library(tibble)
library(gratia)
library(latex2exp)

## set the default theme for ggplot objects to theme_minimal()
theme_set(theme_minimal())
theme_update(panel.grid.major = element_line(size = 0.25),
             panel.grid.minor = element_line(size = 0.15))

# check for the R version number and if less than 3.6, switch to using the
# new random number generator, to ensure replicability

if(getRversion() < 3.6) RNGversion("3.6.0")
```

```{r setup_tables, include=FALSE, purl=FALSE}
#This code is for configuring latex tables. It is not needed for general script
#usage.
table_out_format <- ifelse("pdf_document" %in% rmarkdown::all_output_formats("../paper_sections/full_document.Rmd"),
                    "latex",
                     ifelse("html_document" %in% rmarkdown::all_output_formats("../paper_sections/full_document.Rmd"),
                            "html",
                            NA_character_)
                      )
```
\vspace{-1cm}
\begin{center}
\author{
{\bf Ian D. Jonsen\affmark[1,]\affmark[*], W. James Grecian\affmark[2], Lachlan Phillips\affmark[1], Gemma Carroll\affmark[3], Robert G. Harcourt\affmark[1], and Toby A. Patterson\affmark[4]}\\ \vspace{2mm}
\affaddr{\affmark[1]Department of Biological Sciences, Macquarie University, Sydney, NSW, Australia}\\
\affaddr{\affmark[2]Scottish Oceans Institute, University of St Andrews, KY16 8LB, United Kingdom}\\
\affaddr{\affmark[3]Environmental Defense Fund, Seattle, WA, United States}\\
\affaddr{\affmark[4]CSIRO Ocean and Atmosphere Research, Hobart, TAS, Australia}\\
\affaddr{\affmark[*]corresponding author, ian.jonsen@mq.edu.au}\\
}
\end{center}

# Abstract
1.  

2.  

3.  

4.  

### Keywords: 

```{r child = '01-intro.Rmd'}
```
\FloatBarrier
```{r child = '02-overview.Rmd'}
```
\FloatBarrier
```{r child = '03-examples.Rmd'}
```
\FloatBarrier
```{r child = '04-discussion.Rmd'}
```

# Acknowledgements
We thank xx, xx for helpful comments on an earlier draft of this manuscript. We thank Marie Auger-Méthé for contributing original code to the movement persistence models. IDJ acknowledges support from a Macquarie University co-Funded Fellowship and from partners: the US Office of Naval Research, Marine Mammal Program (grant N00014-18-1-2405); the Integrated Marine Observing System (IMOS); Taronga Conservation Society; the Ocean Tracking Network; Birds Canada; and Innovasea/VEMCO. TAP was supported by CSIRO Oceans & Atmosphere internal research funding scheme.

# Author's Contributions
IDJ developed the R package; WJG contributed harp seal data and to the R package; LP, GC, and RGH contributed little penguin data; IDJ and TAP developed the state-space models; IDJ wrote an initial draft of the manuscript with a contribution from WJG; all authors edited the manuscript.

# Data Accessibility 
All code mentioned here is provided in the `foieGras` package for R available on CRAN at \url{https://CRAN.R-project.org/package=foieGras}. The development version of the package is available on GitHub at \url{https://github.com/ianjonsen/foieGras}. Data used in the examples are available at...

# ORCID
*Ian D Jonsen* \url{https://orcid.org/0000-0001-5423-6076}  
*Toby A Patterson* \url{https://orcid.org/0000-0002-7150-9205}

# References
