---
title: '`foieGras` an R package for animal movement data: rapid quality control, behavioural estimation and simulation'
header-includes:
- \usepackage{lineno}
- \usepackage{placeins}
- \linenumbers
- \usepackage{sectsty}
- \sectionfont{\sffamily\large\scshape}
- \subsectionfont{\sffamily\normalsize\scshape}
compact-title: no
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    citation_package: natbib
    includes:
      in_header: preamble.sty
  html_document:
    df_print: paged
  word_document: default
documentclass: article
classoption: 12pt
geometry: left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm,headheight=12pt,letterpaper
bibliography: references.bib
biblio-style: mee
spacing: double
---

```{r knitr_setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=TRUE,
                      warning=FALSE,
                      cache=TRUE,
                      autodep=TRUE,
                      fig.path="../figures/",
                      #to handle remaining figures without indices
                      #From:https://github.com/yihui/knitr/issues/505
                      fig.process = function(x) {
                        x2 = sub('-\\d+[.][a-z]+)$', '\\1', x)
                        if (file.rename(x, x2)) x2 else x 
                        }
                      )
```

```{r setup, include=FALSE, cache=FALSE}
#### Setup ####
## all the packages needed for this tutorial are listed here
library(foieGras)
library(ggplot2)
library(ggthemes)
library(patchwork)
library(dplyr)
library(lubridate)
library(kableExtra)
library(knitr)
library(docxtools)
library(tibble)
library(latex2exp)

## set the default theme for ggplot objects to theme_minimal()
theme_set(theme_minimal())
theme_update(panel.grid.major = element_line(size = 0.25),
             panel.grid.minor = element_line(size = 0.15))

# check for the R version number and if less than 3.6, switch to using the
# new random number generator, to ensure replicability

if(getRversion() < 3.6) RNGversion("3.6.0")
```

```{r setup_tables, include=FALSE, purl=FALSE}
#This code is for configuring latex tables. It is not needed for general script
#usage.
table_out_format <- ifelse("pdf_document" %in% rmarkdown::all_output_formats("../paper_sections/full_document.Rmd"),
                    "latex",
                     ifelse("html_document" %in% rmarkdown::all_output_formats("../paper_sections/full_document.Rmd"),
                            "html",
                            NA_character_)
                      )
```
\vspace{-1cm}
\begin{center}
\author{
{\bf Ian D. Jonsen\affmark[1,]\affmark[*], W. James Grecian\affmark[2], Lachlan Phillips\affmark[1], Gemma Carroll\affmark[3], Clive McMahon\affmark[4], Robert G. Harcourt\affmark[1], Mark A. Hindell\affmark[5], and Toby A. Patterson\affmark[6]}\\ \vspace{2mm}
\affaddr{\affmark[1]School of Natural Sciences, Macquarie University, Sydney, NSW, Australia}\\
\affaddr{\affmark[2]Sea Mammal Research Unit, Scottish Oceans Institute, University of St Andrews, St Andrews, Fife, United Kingdom}\\
\affaddr{\affmark[3]Environmental Defense Fund, Seattle, WA, United States}\\
\affaddr{\affmark[4]Sydney Institute of Marine Science, Mosman, NSW, Australia}\\
\affaddr{\affmark[5]Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, TAS, Australia}\\
\affaddr{\affmark[6]CSIRO Ocean and Atmosphere Research, Hobart, TAS, Australia}\\
\affaddr{\affmark[*]corresponding author, ian.jonsen@mq.edu.au}\\
}
\end{center}

# Abstract
1.  Animal-borne sensor and tracking data are essential for understanding the ecology, behaviour and physiology of mobile or cryptic species. Biologically meaningful signals in these data can be obscured by noise due to imperfect measurement technology, and rigorous quality control is required before before comprehensive analysis can proceed.

2.  State-space models are powerful tools that can separate artificial noise from biologically meaningful signals. These tools are ideal for quality control of error-prone location data and for inferring where animals are and what they are doing when they record or transmit sensor information. However, these are statistically complex models that can be challenging to fit to diverse animal tracking data sets.

3.  We introduce the R package `foieGras` that aims to 

4.  

### Keywords: 

```{r child = '01-intro.Rmd'}
```
\FloatBarrier
```{r child = '02-overview.Rmd'}
```
\FloatBarrier
```{r child = '03-applications.Rmd'}
```
\FloatBarrier
```{r child = '04-conclusions.Rmd'}
```

# Acknowledgements
We thank Marie Auger-Méthé for contributing original code to the movement persistence models. IDJ acknowledges support from a Macquarie University co-Funded Fellowship and from partners: the US Office of Naval Research, Marine Mammal Program (grant N00014-18-1-2405); the Integrated Marine Observing System (IMOS); Taronga Conservation Society; the Ocean Tracking Network; Birds Canada; and Innovasea/VEMCO. TAP was supported by CSIRO Oceans & Atmosphere internal research funding scheme. The Integrated Marine Observing System (IMOS) supported seal fieldwork. IMOS is a national collaborative research infrastructure, supported by the Australian Government and operated by a consortium of institutions as an unincorporated joint venture, with the University of Tasmania as Lead Agent. Field work at Illes Kerguelen was conducted as part of the IPEV programme N^o^ 109 (PI H. WEIMERSKIRCH) and of the SNO-MEMO programme (PI C. GUINET) in collaboration with IMOS. CTD tags were partly funded by CNES-TOSCA and IMOS. Little penguin fieldwork was supported by an Australian Research Council Linkage grant to IDJ, GC and RGH (LP160100162). Harp seal fieldwork was funded by Fisheries and Oceans Canada, a bursary from Department for Business, Energy and Industrial Strategy (BEIS) administered by the NERC Arctic Office, and the ARISE project (NE/P006035/1 and NE/P00623X/1), part of the Changing Arctic Ocean programme jointly funded by the UKRI Natural Environment Research Council (NERC) and the German Federal Ministry of Education and Research (BMBF). All animal tagging procedures approved and executed under the Animal Ethics Committee guidelines of the University of Tasmania (elephant seals), Macquarie University (little penguins), and University of St Andrews (harp seals).

# Author's Contributions
IDJ developed the R package; WJG contributed harp seal data and to the R package; LP, GC, and RGH contributed little penguin data; CRM, RGH and MAH contributed Southern elephant seal data; IDJ and TAP developed the state-space models; IDJ wrote an initial draft of the manuscript with contributions from WJG; all authors contributed to the final manuscript.

# Data Accessibility 
All code and data used here are provided in the `foieGras` package for R or in the Supplementary Information. The latest stable and cross-platform tested version of the package (currently, 1.0-7) is available via ROpenSci's R-universe, at https://ianjonsen.r-universe.dev/ui#package:foieGras. The latest partially tested stable and development versions are available on the GitHub repository: https://github.com/ianjonsen/foieGras. An older version of `foieGras` (0.7-6) remains on CRAN at \url{https://CRAN.R-project.org/package=foieGras}, however, we recommend users upgrade to the latest R-universe version for full access to the functionality presented here. 
 
# ORCID
*Ian D Jonsen* \url{https://orcid.org/0000-0001-5423-6076}  
*W James Grecian* \url{https://orcid.org/0000-0002-6428-719X}  
*Lachlan Phillips* \url{https://orcid.org/0000-0002-7635-2817}  
*Gemma Carroll* \url{https://orcid.org/0000-0001-7776-0946}  
*Clive R McMahon* \url{https://orcid.org/0000-0001-5241-8917}  
*Robert G Harcourt* \url{https://orcid.org/0000-0003-4666-2934}  
*Mark A Hindell* \url{https://orcid.org/0000-0002-7823-7185}  
*Toby A Patterson* \url{https://orcid.org/0000-0002-7150-9205}  

# References
